---
title: "PracticalMachineLearning"
author: "Sasa Pakvovic"
date: "Saturday, December 20, 2014"
output: html_document
---

#Background

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website [here:](http://groupware.les.inf.puc-rio.br/har) (see the section on the Weight Lifting Exercise Dataset). 

###Data
The data comes form this [source](http://groupware.les.inf.puc-rio.br/har).

###Analysis
We will be using data acquired to build up a model to be able to predict how well 
the practicioner is doing the exercise. For this purpose we will be creating random forest model because of its robustness in many situations.

###Data sliceing
We will be using some of the available R packages to prepare the model.

```{r}
library(caret)
library(kernlab)
library(randomForest)
library(corrplot)
set.seed(1050)
```

```{r,echo=FALSE,cache=TRUE}
# check if a data folder exists; if not then create one
if (!file.exists("data")) {dir.create("data")}

# download files
filesource <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
destfile1 <- "./data/trainpml"
download.file(filesource,destfile1,method="curl")

filesource <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
destfile2 <- "./data/testpml"
download.file(filesource,destfile2,method="curl")

# read the csv file
train <- read.csv("./data/trainpml", na.strings= c("NA",""," "))
```

After loading and looking a little bit at the data using eploratory data analysis
very soon is clear that data has a lot of NAs and has columns which can not help
us in building a model(metadata).

```{r}
# cleaning of the data needed
trainNAs <- apply(train, 2, function(x) {sum(is.na(x))})
trainclean <- train[,which(trainNAs==0)]

# for prediction we will not include firsr 8 rows
trainclean <- trainclean[8:length(trainclean)]

```

###Model building

On clean data we need to divide the train data set into training and cross-validation datasets. Iv decided to take a 80/20 split.

```{r}
# split the train data into 2 sets - training and cross validation
inTrain <- createDataPartition(y = trainclean$classe, p = 0.8, list = FALSE)
training <- trainclean[inTrain, ]
cval <- trainclean[-inTrain, ]
```

We can take a look at the correlation matrix of all the predictors.

```{r}
# calc the correlation matrix
cormatrix <- cor(training[, -length(training)])
corrplot(cormatrix, order = "FPC", method = "circle", type = "lower", tl.cex = 0.7,  tl.col = rgb(0, 0, 0))
```

We are going to deploy a radnom forest model to predict classe using all other 
feautres as predictors. We will use crossvalidation to build up a confusion matrix and check how the model is performing.

```{r}
# fit a radnom forest model 
modelFit <- randomForest(classe ~ ., data = training)
# crossvalidate the model
predictCVal <- predict(modelFit, cval)
confusionMatrix(cval$classe, predictCVal)
```

As we can see the model is performing at an accuracy of 99,7%, so the out-of-sample error is 0,03%.

###Using the model on the test dataset

Only thing left to do is use our model on the the test dataset and predict the 
classes of classe.

```{r}
# now use the model on the test data
test <- read.csv("./data/testpml", na.strings= c("NA",""," "))
testNAs <- apply(test, 2, function(x) {sum(is.na(x))})
testclean <- test[,which(testNAs==0)]
testclean <- testclean[8:length(testclean)]
predictTest <- predict(modelFit, testclean)
predictTest
```


